---
title: "Zhong_expression_file"
author: "Gitte Aasbjerg"
date: "05/11/2020"
output: html_document
---

# Load data
```{r}
library("Seurat")
Zhong.data <- Read10X(data.dir = "/home/cbmr/kzd307/gitte/hippocampus/data/Zhong_expression_data/")
```


#-------------- Remove cells expressing haemoglobin genes --------------


# Run SCTransform
```{r}
SCTdata <- CreateSeuratObject(counts=Zhong.data)
SCTdata <- PercentageFeatureSet(SCTdata, pattern = "^MT-", col.name = "percent.mt")
SCTdata <- SCTransform(SCTdata, vars.to.regress = "percent.mt", verbose = FALSE)
```

# Run PCA and UMAP to cluster data
```{r}
SCTdata <- RunPCA(SCTdata, verbose = FALSE)
SCTdata <- RunUMAP(SCTdata, dims = 1:15, verbose = FALSE)

SCTdata <- FindNeighbors(SCTdata, dims = 1:15, verbose = FALSE)
SCTdata <- FindClusters(SCTdata, verbose = FALSE)

DimPlot(SCTdata, label = TRUE) + NoLegend()
```

# Locate cluster with cells expressing haemolglobin
```{r}
haemoglobin <- FeaturePlot(SCTdata, features = c('HBM', 'HBA1', 'HBA2', 'HBB', 'HBD', 'HBE1', 'HBG1', 'HBG2', 'HBQ1', 'HBZ'), pt.size = 0.2, ncol = 3)
haemoglobin
```

# Identify cells from cluster 12 and 13 
```{r}
remove_clusters <- which(SCTdata@meta.data$seurat_clusters == 13 | SCTdata@meta.data$seurat_clusters == 12) 
all.cells <- colnames(SCTdata)
remove_cells <- all.cells[remove_clusters]
```



#--------------- Initiate original pre-processing steps --------------------


# Read 10X data and create Seurat Object and exclude cells expressing haemoglobin genes
```{r}
data <- CreateSeuratObject(counts=Zhong.data)
data <- data[,!colnames(data) %in% remove_cells]
```

# Perform SCTransform of reduced data and remove cells expressing mitochondrial genes
```{r}
data <- PercentageFeatureSet(data, pattern = "^MT-", col.name = "percent.mt")
data <- SCTransform(data, vars.to.regress = "percent.mt", verbose = FALSE)
```

# Evalutate number of dimensions in the data
```{r}
ElbowPlot(data)
DimHeatmap(data, dims = 1:15, cells = 500, balanced = TRUE)
```

# Dimensions is set to be 13 based on the above results. The data is then clustered
```{r}
data <- RunPCA(data, verbose = FALSE)
data <- RunUMAP(data, dims = 1:13, verbose = FALSE)

data <- FindNeighbors(data, dims = 1:13, verbose = FALSE)
data <- FindClusters(data, verbose = FALSE)

DimPlot(data, label = TRUE) + NoLegend()
```

# Save expression file and cluster annotations
```{r}

# Expression matrix
data_matrix <- as.matrix(GetAssayData(object = data, slot = "counts"))
write.csv(data_matrix, file="../data/Zhong_expression_matrix.csv")


# Cluster annotations
gene_list <- as.matrix(colnames(data))
cluster <- as.matrix(as.numeric(data@meta.data$seurat_clusters))
cluster_annotation = cbind(gene_list, cluster)

# Write cluster annotation file
write.csv(cluster_annotation, file = "Zhong_cluster_annotation.csv")

saveRDS(data, file = "../data/Zhong_SCTransformed.rds")

```

